### 2017-05-10 manuscript/phylotyper_suppl_info.tex

\documentclass[11pt,letterpaper]{article}
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{ragged2e,booktabs,multirow}
\usepackage{caption,capt-of}
\usepackage{graphicx}
\usepackage{array}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\newcolumntype{M}[1]{>{\begin{varwidth}[t]{#1}}l<{\end{varwidth}}}

\RequirePackage[LY1]{fontenc}

\renewcommand{\rmdefault}{ptm}
\renewcommand{\sfdefault}{phb}
\renewcommand{\ttdefault}{pcr}

%\RequirePackage[LY1,mtbold]{mathtime}
\def\helvetica{\fontfamily{phv}\selectfont}
\def\helveticaitalic{\fontfamily{phv}\itshape\selectfont}
\def\helveticabold{\fontfamily{phv}\bfseries\selectfont}
\def\helveticabolditalic{\fontfamily{phv}\bfseries\itshape\selectfont}
\def\helveticacn{\fontfamily{phv}\fontseries{mc}\fontshape{n}\selectfont}
\def\helveticacnitalic{\fontfamily{phv}\fontseries{mc}\fontshape{sl}\selectfont}
\def\helveticacnbold{\fontfamily{phv}\fontseries{bc}\fontshape{n}\selectfont}
\def\helveticacnbolditalic{\fontfamily{phv}\fontseries{bc}\fontshape{sl}\selectfont}

\title{Phylotyper: {\it In silico} predictor of molecular subtypes from gene sequences \\ \large Supplementary Information}
\author{Matthew D. Whiteside, Chad R. Laing and Victor P.J. Gannon}
\date{ }

\begin{document}

%\maketitle

\tableofcontents

\clearpage

\section{Supplementary Methods}

\subsection{Performance Analysis}

\subsubsection{BLAST-based method}

BLAST

\subsubsection{Selecting Percent Identity Cutoff for BLAST Approach}

Following \citep{Jenkins2015}, percent identity in the BLAST alignments is used as a cutoff. Subtype classifications are transfered from the top BLAST match provided the alignment percent identity is above the cutoff. The cutoff impacts performance in one of two ways: too low and non-equivalent subtype alleles may be matched with the query (e.g. type I error), too high and valid subtype allele BLAST hits are filtered out (e.g. type II error). To select the ideal cutoff that balances these competing sources of error, a cross-validation simulation was run to estimate false discovery and true positive rates across varying percent identity cutoffs. A 10-fold cross-validation was performed; 10\% of the training set was randomly selected. The remaining 90\% was used to build the reference BLAST database. The 10\% test set was then searched against this database using BLAST and positive and negative results were recorded. This step was repeated 100 times and the F1-score, which is a weighted average of the precision and recall, was calculated. Precision is defined as:

$Precision = TP / (TP + FP)$

and recall as:

$Recall = TP / (TP + FN)$

These values are then combined to produce an F-score as follows:

$2 * ( precision * recall ) / ( precision + recall )$

The F-score; a binary classifer metric, was recorded individually for each subtype and then averaged to create a single performance metric for this multi-class predictor.  F-score was computed for increasing percent identities, and the percent identity corresponding to the maximum F-score was selected as the cutoff.

\subsubsection{Performance Assessment}

Two separate analyses

\section{Supplementary Figures}

\subsection{Supplementary Figure S1}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{sfig01.png}
\caption{Stx2 Phylogenetic tree showing the Phylotyper marginal likelihoods as pie charts. This output is provided to the user.}
\end{figure}

\clearpage

\section{Supplementary Tables}

\subsection{Supplementary Table S1}

~

\begin{minipage}{\linewidth}
\centering
\captionof{table}{Subtype Schemes in Phylotyper}
\medskip
\begin{tabular}{@{}llll@{}}\toprule Name &
Description & Species & Loci\\\midrule
stx1 & Shiga-toxin 1 subtype & {\it E. coli} & 2 \\
stx2 & Shiga-toxin 1 subtype & {\it E. coli} & 2\\
eae & Intimin subtype & {\it E. coli} & 1\\
flic & H-serotype based on flagellin gene; fliC & {\it E. coli} & 1\\
wz & O-serotype based on the wzy and wzx genes & {\it E. coli} & 2\\\bottomrule
\end{tabular}\par

\end{minipage}

\clearpage

\subsection{Supplementary Table S2}

Table S2 contains performance metrics from a leave-one-out cross-validation analysis comparing Phylotyper and a top-BLAST hit approach.  The analysis examines the four \textit{E. coli} schemes available in Phylotyper. In this multi-class analysis, precision, recall and F$_{1}$ score are calculated for each individual class provided that at least one instance of the class is in the training set.  The individual class positive and negatives are summed to calculate an overall precision, recall and F$_{1}$ score for the scheme.

%\captionsetup{justification=RaggedRight}

\begin{minipage}{\linewidth}
\centering
\setlength{\tabcolsep}{4pt}
\captionof{table}[justification=RaggedRight]{Leave-One-Out Cross Validation Results}
\medskip
\begin{tabular}{@{\extracolsep{4pt}}llll>{\centering}m{1.6cm}lll@{}}
\toprule 
\multirow{2}{*}{Scheme} & \multicolumn{4}{c}{Phylotyper} & \multicolumn{3}{c}{Sequence-similarity}\\
\cline{2-5}\cline{6-8}
& Precision & Recall & F$_{1}$ Score & Run-time (s) $^{1}$ & Precision & Recall & F$_{1}$ Score \\
\midrule
{\it E. coli} Stx1 & 1.00 & 0.94 & 0.97 & 6 & 0.94 & 0.94 & 0.94\\
{\it E. coli} Stx2 & 1.00 & 0.99 & 0.99 & 32 & 0.93 & 0.93 & 0.93\\
{\it E. coli} Intimin & 1.00 & 0.98 & 0.99 & 17 & 0.99 & 0.98 & 0.99\\
{\it E. coli} H-serotype & 0.99 & 0.98 & 0.98 & 16 & 0.96 & 0.85 & 0.90\\
{\it E. coli} O-serotype & 1.00 & 0.61 & 0.75 & 67 & 1.00 & 0.36 & 0.53\\\bottomrule
\end{tabular}\par
\bigskip
\raggedright
Formula:
\begin{enumerate}
\item $Precision = TP / (TP + FP)$
\item $Recall = TP / (TP + FN)$
\item $F_{1}~score = 2*Precision*Recall/(Precision + Recall)$
\end{enumerate}
$TP = True~Positive,~~FP = False~Positive,~~FN = False~Negative$
\par
\bigskip
$^{1}$ Run-time is for an {\it Escherichia coli} genome containing a non-identical subtype gene which triggers the full phylotyper pipeline.

\end{minipage}

\end{document}