### 2017-05-16 manuscript/phylotyper_suppl_info_revised.tex
UPDATE: intro UPDATE: implementation UPDATE: suppl info UPDATE: response to reviewers

\documentclass[11pt,letterpaper]{article}
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{ragged2e,booktabs,multirow}
\usepackage{caption,capt-of}
\usepackage{graphicx}
\usepackage{array}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\newcolumntype{M}[1]{>{\begin{varwidth}[t]{#1}}l<{\end{varwidth}}}

\RequirePackage[LY1]{fontenc}

\renewcommand{\rmdefault}{ptm}
\renewcommand{\sfdefault}{phb}
\renewcommand{\ttdefault}{pcr}

%\RequirePackage[LY1,mtbold]{mathtime}
\def\helvetica{\fontfamily{phv}\selectfont}
\def\helveticaitalic{\fontfamily{phv}\itshape\selectfont}
\def\helveticabold{\fontfamily{phv}\bfseries\selectfont}
\def\helveticabolditalic{\fontfamily{phv}\bfseries\itshape\selectfont}
\def\helveticacn{\fontfamily{phv}\fontseries{mc}\fontshape{n}\selectfont}
\def\helveticacnitalic{\fontfamily{phv}\fontseries{mc}\fontshape{sl}\selectfont}
\def\helveticacnbold{\fontfamily{phv}\fontseries{bc}\fontshape{n}\selectfont}
\def\helveticacnbolditalic{\fontfamily{phv}\fontseries{bc}\fontshape{sl}\selectfont}

\title{Phylotyper: {\it In silico} predictor of molecular subtypes from gene sequences \\ \large Supplementary Information}
\author{Matthew D. Whiteside, Chad R. Laing and Victor P.J. Gannon}
\date{ }

\begin{document}

%\maketitle

\tableofcontents

\clearpage

\section{Supplementary Methods}

\subsection{Phylotyper Verifications Performed on New Subtype Schemes}

Phylotyper conducts two tests for submitted new subtype schemes. The purpose of the first check is to identify isolated cases of distinct subtypes that are clustered together in the phylogenetic tree. In large-scale cases, descendent clades in the phylogenetic tree with multiple distinct subtypes may indicate a subtype that is not correlated with the phylogeny. In limited cases, it might indicate annotation errors. Phylotyper computes an inter-patritristic distances for all leaves in the subtype phylogenetic tree for the given subtype scheme. It then computes a threshold is equivalent to 0.4 probability that inter-patristic distances equal to or greater then threshold belong to the same-subtype distribution. To estimate the distributions, all inter-patristic distances for nodes with the same subtype and all distances for nodes with different subtypes are fit with R's Mclust package. Subtrees with a max inter-patristic distance less than this threshold that have distinct subtypes are flagged. Users will be notified and an updated subtype scheme will be generated with proposed corrections when this situation is detected.

All new subtype schemes are subject to a leave-one-out cross-validation analysis. You will be notified if the F1-score (a equally weighted average of precision and recall) is below 0.9.

\subsection{BLAST-based method}

For comparison purposes, we developed a generalized sequence-similarity predictor of subtypes using BLAST. Following \citep{Jenkins2015}, a training set of sequences with annotated subtypes is used to build a BLAST database (the training sets are from the the Phylotyper repository). To predict a subtype, query sequences are searched against the training BLAST database and the top hit is recorded. If the top hit's alignment with the query has at least 60\% coverage with the reference sequence and is above the percent identity cutoff, then the subtype of the top hit is returned as the predicted subtype of the query sequence. The cutoff impacts performance in one of two ways: too low and non-equivalent subtype alleles may be matched with the query (e.g. type I error), too high and valid subtype allele BLAST hits are filtered out (e.g. type II error). To select the ideal cutoff that balances these competing sources of error, a cross-validation simulation was run to estimate false discovery and true positive rates across varying percent identity cutoffs. A 10-fold cross-validation was performed; 10\% of the training set was randomly selected. The remaining 90\% was used to build the reference BLAST database. The 10\% test set was then searched against this database using BLAST and positive and negative results were recorded. This step was repeated 100 times and the F1-score, which is a weighted average of the precision and recall, was calculated. Precision is defined as:

$Precision = TP / (TP + FP)$

and recall as:

$Recall = TP / (TP + FN)$

These values are then combined to produce an F-score as follows:

$2 * ( precision * recall ) / ( precision + recall )$

The F-score; a binary classifer metric, was recorded individually for each subtype and then averaged to create a single performance metric for this multi-class predictor.  F-score was computed for increasing percent identities, and the lowest percent identity corresponding to the maximum F-score was selected as the cutoff. The percent identity cutoff was selected as 0.97.

\subsection{Performance Assessment}

Two separate cross-validation tests were performed to compare the performance of Phylotyper with the BLAST method. The comparison to vtyper is a separate analysis, as vtyper does not use a training set and a cross-validation test would not work in this situation. A leave-one-out cross-validation analysis was performed where each sequence in the training set is withheld and the remaining sequences are used for form the phylogenetic tree for Phylotyper or are used to build the BLAST database for the BLAST-based approach. The predicted result for the test sequence is then compared to the true value.

The second validation iterated over each subtype and removed all genes assigned that subtype from the training set. The removed genes or test set was then used as input to identify the level of false positive predictions that occur when the tools are presented with novel subtypes not in the training set.

Precision, recall and the F-score as defined earlier for a multi-classifier test, were computed for the validations.

\subsubsection{V-typer Assessment}

The V-typer tool predicts Stx1 or Stx2 subtypes by simulating PCR (i.e. mapping primers). Due to the unique method in V-typer,  which does not use a training dataset, the cross-validation assessment was adapted to test V-typer. Each experimentally-verified Stx gene sequence that makes up Phylotyper training dataset was searched for against the NCBI nt database to retrieve surrounding sequence that may be involved in the PCR detection by V-typer.  Specifically, regions that have at least 5 kB up- and downstream from the Stx gene were selected for the assessment. V-typer was run on each sequence and the results recorded in supplmentary table XXXX.

\section{Supplementary Figures}

\subsection{Supplementary Figure S1}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{sfig01.png}
\caption{Stx2 Phylogenetic tree showing the Phylotyper marginal likelihoods as pie charts. This output is provided to the user.}
\end{figure}

\clearpage

\subsection{Supplementary Figure S2}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{sfig02.png}
\caption{The effect of randomizing a specified proportion of subtype labels in the Eae phylogenetic tree on the empirically estimated performance metrics: F$_1$-score and recall.}
\end{figure}

\clearpage

\section{Supplementary Tables}

\subsection{Supplementary Table S1}

~

\begin{minipage}{\linewidth}
\centering
\captionof{table}{Subtype Schemes in Phylotyper}
\medskip
\begin{tabular}{@{}llll@{}}\toprule Name &
Description & Species & Loci\\\midrule
stx1 & Shiga-toxin 1 subtype & {\it E. coli} & 2 \\
stx2 & Shiga-toxin 1 subtype & {\it E. coli} & 2\\
eae & Intimin subtype & {\it E. coli} & 1\\
flic & H-serotype based on flagellin gene; fliC & {\it E. coli} & 1\\
wz & O-serotype based on the wzy and wzx genes & {\it E. coli} & 2\\\bottomrule
\end{tabular}\par

\end{minipage}

\clearpage

\subsection{Supplementary Table S2}

Table S2 contains performance metrics from a leave-one-out cross-validation analysis comparing Phylotyper and a top-BLAST hit approach.  The analysis examines the four \textit{E. coli} schemes available in Phylotyper. In this multi-class analysis, precision, recall and F$_{1}$ score are calculated for each individual class provided that at least one instance of the class is in the training set.  The individual class positive and negatives are summed to calculate an overall precision, recall and F$_{1}$ score for the scheme.

%\captionsetup{justification=RaggedRight}

\begin{minipage}{\linewidth}
\centering
\setlength{\tabcolsep}{4pt}
\captionof{table}[justification=RaggedRight]{Leave-One-Out Cross Validation Results}
\medskip
\begin{tabular}{@{\extracolsep{4pt}}llll>{\centering}m{1.6cm}lll@{}}
\toprule 
\multirow{2}{*}{Scheme} & \multicolumn{4}{c}{Phylotyper} & \multicolumn{3}{c}{Sequence-similarity}\\
\cline{2-5}\cline{6-8}
& Precision & Recall & F$_{1}$ Score & Run-time (s) $^{1}$ & Precision & Recall & F$_{1}$ Score \\
\midrule
{\it E. coli} Stx1 & 1.00 & 0.94 & 0.97 & 6 & 0.94 & 0.94 & 0.94\\
{\it E. coli} Stx2 & 1.00 & 0.99 & 0.99 & 32 & 0.93 & 0.93 & 0.93\\
{\it E. coli} Intimin & 1.00 & 0.98 & 0.99 & 17 & 0.99 & 0.98 & 0.99\\
{\it E. coli} H-serotype & 0.99 & 0.98 & 0.98 & 16 & 0.96 & 0.85 & 0.90\\
{\it E. coli} O-serotype & 1.00 & 0.61 & 0.75 & 67 & 1.00 & 0.36 & 0.53\\\bottomrule
\end{tabular}\par
\bigskip
\raggedright
Formula:
\begin{enumerate}
\item $Precision = TP / (TP + FP)$
\item $Recall = TP / (TP + FN)$
\item $F_{1}~score = 2*Precision*Recall/(Precision + Recall)$
\end{enumerate}
$TP = True~Positive,~~FP = False~Positive,~~FN = False~Negative$
\par
\bigskip
$^{1}$ Run-time is for an {\it Escherichia coli} genome containing a unique gene which triggers the full phylotyper pipeline.

\end{minipage}

\begin{minipage}{\linewidth}
\centering
\captionof{table}{V-typer Performance}
\medskip
\begin{tabular}{@{}llll@{}}\toprule
NCBI Accession & Stx Subtype & V-typer Prediciton$^{1}$ & Phylotyper Prediction\\\midrule
CP015020 & 2d & N & 2d \\
CP015229 & 2g & 2g & 2g \\
CP009106 & 2a & N & 2a \\
CP007133 & 2a & 2a & 2a \\
CP006262 & 2a & 2a & 2a \\
AP010960 & 2a & 2a & 2a \\
CP013663 & 2a & N & 2a \\
CP015228 & 2a & N & 2a \\
CP011331 & 2a & N & 2a \\
HF572917 & 2a & N & 2a \\
LN554923 & 2a & N & 2a \\
LM997071 & 2a & N & 2a \\
CP018250 & 2c & N & 2c \\
CP018252 & 2c & N & 2c \\
CP018247 & 2c & N & 2c \\
CP018245 & 2c & N & 2c \\
CP018243 & 2c & N & 2c \\
CP017446 & 2c & N & 2c \\
CP017442 & 2c & N & 2c \\
CP017438 & 2c & N & 2c \\
CP014314 & 2c & N & 2c \\
LM997036 & 2a & 2a & 2a \\
LM996489 & 1c & 1c & 1c \\
LM997036 & 1c & 1c & 1c \\
Total & 24 & 7 & 24 \\
\bottomrule
\end{tabular}\par
\bigskip
\raggedright
$^{1}$ N=Negative (no results returned)

\end{minipage}

\end{document}