### 2017-05-30 manuscript/response_to_reviewers.txt
EDIT: finished response to reviewers

response_to_reviewers.txt

Firstly, we would like to thank the reviewers for their thoughtful comments. We have made subtantual changes to the results and discussion sections as well as to the Phylotyper code that we hope address your suggestions for the Phylotyper manuscript.  We have responded to each individual comment below:

----------------------------------------

Here are the comments of the reviewers:

----------------------------------------
Reviewer: 1

Comments to the Author
This is a great little tool that I can see being integrated into numerous pipelines. It moves away from the conventional MLST scheme and is a little more precise than the large scale core genome typing methods being used. I was able to successfully download, install and run the application. I also found the output files very useful and clear, something that doesn't happen enough with many bioinformatics applications.

I have two minor comments that are more "food for thought" than required revisions. I am good with the note as is, everything is clear, well cited and concise. For more information it is possible to visit the GitHub page.

"Food for thought" comments:

Although the GitHub page is well laid out and easy enough to follow, please consider including some sort of example to follow along with. As a newcomer to the application it took me a little while to figure out how to get everything exactly right for an initial run. I think a very small example to follow and run that is located early in the GitHub instructions would really help to just get things going.

*Response:
We agree with the reviewer's suggestion that a working example for new users will enhance the adoption and usability of Phylotyper.  We have added an "Example" section to the README.md file (which is also the landing page for Phylotyper's GitHub repository: https://github.com/superphy/insilico-subtyping/#examples) along with example data that demonstrates how to 1) predict subtypes for schemes already packaged in Phylotyper and 2) how to add and run user-defined subtype schemes.

There lies a slight issue in the "user generated sub-type schemes" in that the results that you get with Phylotyper are highly dependant on the quality of the sub-type scheme provided. Since this tool will likely be used in outbreak situations and analyses, do you have any plans to address the quality of the sub-type scheme that was used? Do you plan on having a section in the GitHub page that allows users to send new sub-type schemes for validation and posting?

*Response*:
The reviewer is correct that Phylotyper's performance is dependent on the input reference sequences used to build the subtype predictor. To mitigate poor prediction from unsuitable or inaccurate inputs, we included in the build pipeline for the local user-submitted subtypes, a series of checks to automatically test the performance of any submitted subtype scheme. These checks are now described in the Implementation section.

While these checks identify major breakdowns in the training set used by Phylotyper, we note there is no check to identify missing data, specifically, missing subtypes or key subtypes gene sequences that define the range of the subtype sequence space. For any current general subtype predictor (i.e. not predictors that replicate PCR in silico, for instance), performance is dependent on the reference input set. In dealing with missing data, we demonstrate that Phylotyper is significantly more robust than a direct sequence similarity-based predictor. It is less likely to generate a type-I error when encountering a subtype not found in the training set. The Phylotyper method estimates the uncertainty associated with a subtype prediction. When test sequences that are phylogenetically distant from known reference sequences or in a region of the phylogenetic tree that is rapidly evolving, the Phylotyper score will reflect the degree of uncertaintity. Most often the sequence for a novel or missing subtype will be relatively phylogenetically distant from the other subtype sequences in the tree and its Phylotyper uncertaintity score will reflect this. Only in cases where the novel/missing subtype is closely phylogenetically related to an existing subtype in the training set, will Phylotyper generate a false positive prediction. This property of Phylotyper makes it more robust when working with incomplete training data.

To enhance the users understanding of the checks that are provided in the build feature and the potential blindspots in creating their own subtype predictors with Phylotyper, we have greatly expanded the methods section and the README provided with the software. The added content describes the two checks, how they work and what they detect. In manuscript we also provide guidelines about the inputs needed for an effective subtype predictor. We also included in the manuscript a new analysis that demonstrates how Phylotyper versus a BLAST-based predictor performs when an entire subtype class is removed from the trainging set and used as test input. This analysis and results have been added to the Results and Discussions sections.

The reviewer suggests having a means for users to submit subtypes for evaluation and posting to a shared repository. We agree that this would be a helpful system. It will allow microbiologist specialists to submit and improve subtype schemes and increase coverage. We mention in the manscript and on the README, that users can contact us and submit new or modified subtype schemes. We will check the submitted schemes and add them to our Github repository for general use.

Reviewer: 2

Comments to the Author
General comments:
Whiteside et al present here a novel python package for the prediction of bacterial molecular subtypes that are widely used in public health surveillance and outbreak analysis. The release of such a bioinformatic tools is highly valuable and timely with the democratization of WGS in microbiology laboratories worldwide. The article would benefit from some rewriting to provide more details to the reader regarding the use of Phylotyper and how to extend Phylotyper use to other bacteria than Escherichia coli.  Furthermore, a comparison of Phylotyper performance with other existing tools that the authors mention in the introduction would be welcome to assess the strength and weaknesses of this interesting phylogenetic approach.

Major comments:
Introduction:
1.      The authors should elaborate on the "current subtype systems" for readers that would not be familiar with clinical microbiology standards.

*Response*:
We added a high-level description of mechanisms currently in use to perform subtyping, namely phenotype- and DNA-based methods, with specific examples in each category. We feel this will give the readers an understanding of the breadth of subtyping methods. We also describe the characteristics of the subtypes that have been adapted for in silico prediction using WGS data. Our intention was to convey the wide range of subtyping approaches, and how certain subtyping approaches can be recapitulated from WGS data.

Implementation:
2.      Phylotyper currently only has a scheme for Escherichia coli, which limits its use, and the interest resides in the possibility to build schemes for other bacteria. Therefore, it would be useful for the reader to have more information regarding how to build a new scheme in the contribution, and for example which existing resources of typing schemes could be used for main pathogens.

*Response*:
A detailed step-by-step account of how to run Phylotyper's automated build pipeline is provided in the Phylotyper package instructions. We added a reference and link to these instructions in the Implementation section. We also added to the Implmentation section a description of the requirements in Phylotyper for new subtype schemes as well as the checks that the pipeline automatically performs to validate subtypes schemes created by users. In the Discussion section, we elaborate on the ideal sequence regions that should be targeted by users when developing a WGS-based predictor of subtypes with Phylotyper.  Also in the Discussion seciton, we name some current examples of subtyping schemes in other pathogens outside of E. coli that can be predicted with Phylotyper. We did not systematically go through all main pathogens to identify the subtyping schemes that can predicted with Phylotyper. As we state in our revised Discussion, the subtype schemes that can be linked biologically to a specific genome sequence region are the best candidates for WGS-based predictors. To compile a complete list of these type of subtypes requires domain knowledge in the pathogen (Our lab works with E. coli, hence the the initial focus in Phylotyper's resources on E. coli). We did identify and mention select examples in diverse pathogens to show that Phylotyper is not limited to E. coli. Phylotyper can work with any suitable subtype; identifying a suitable subtype, however, requires knowledge of the subtype and its underlying biological mechanism. To help expand Phylotyper to other pathogens, we added a request in the manuscript and on the Phylotyper website for Microbiologists to send us either subtype schemes or additions/corrections to existing subtype schemes.  Subtype schemes that we recieve we will be packaged into the Phylotyper repository for download and general use.

3.      Please provide more details regarding the “assumptions of the Phylotyper approach” and how/why they would be violated when building new subtype schemes.

*Response*: 
The Phylotyper assumptions are now outlined in the Implementation section. The revised Implementation section covers the basis of the assumptions, how they can be violated, and two built-in mechanisms we developed to alert users of possible violations.

Results:
4.      Whiteside et al compare here their tool with another BLAST based approach that they have developed for the comparison, and not methods published by other groups. The description of the method used for both the BLAST method and the cross validation is imprecise and would benefit from more being more detailed within the main article rather than in the supplementary material. Moreover, a comparison - at least theoretical - with other existing tools would strengthen the uniqueness of Phylotyper highlighted in the discussion.

*Response*:
Our main goal in the performance assessment was to compare how the prediction methods generalize to an independent dataset. There may exists differences in the reference training sets that underpin the tools, and these differences may mask the true performance of the method and we wanted to evalute the predictive capability of the methods themselves. SeroTypeFinder is provided as a web service. The software for SeroTypeFinder has not been made publicly available. Without access to the software, we cannot manipulate the training and test datasets to perform cross-validation analysis. The EcOH tool works directly on unassembled small reads and not assembled genome sequences. A direct comparison of the performance with these tools with Phylotyper is not possible. The intention of using a general BLAST-based approach is demonstrate the benefit of Hidden State Prediction with a phylogenetic tree over approaches that directly match input sequences against a reference database. BLAST, as used in our assessment with a percent identity and coverage cutoff, is the basis of the SeroTypeFinder tool. EcOH employs a more sophisticated algorithm for matching alleles in order to filter out spurious small reads that might align by chance. At its core, it is still matching alleles and does not develop a model of subtype/allele sequence evolution to define boundaries on the subtypes. We feel that our performance comparison to the generalized BLAST-based method does make a valid case for using HSP in subtype prediction. We did add a direct comparison of Phylotyper to Vtyper for Stx genes. The performance results are summerized in the Results section.

To address the reviwers comment that the BLAST and cross-validation method descriptions are imprecise, we have added more thorough descriptions of the BLAST method and the cross-validation analysis in the manuscript. Our aim with the revised descriptions is to provide readers sufficient information give them a complete understanding of the mechanics of the methods. Implementation details, such as percent identity cutoff or number of iterations, remain in Supplementary Information. This was done to fit with page constraints of the Application Note format.

As suggested, we also added a theoretical comparison of the tools to the Discussion; the theoretical comparison looks at use cases, potential issues, and additional benefits outside of performance for the tools.

Discussion:
5.      There is no data to support the claim that Phylotyper easily identifies novel subtype. Did the author try to remove all alleles of one subtype and compute the predictions to see if/how phylotyper creates a consistent subtype? Also, when a new subtype is identified, does Phylotyper includes it in his typing scheme automatically? What would be the recommended tests/further analysis recommended to researchers using Phylotyper in the case a new subtype is identified?

*Response*:
The point we were trying make was that more frequently, Phylotyper correctly handles novel subtypes than existing methods. In our emperical testing we found Phylotyper makes fewer Type-I errors when encountering a novel allele of an existing subtype or a allele with a entirely new subtype not found in reference training set. In former case, Phylotyper returns non-significant/undetermined. We have clarified this point throughout the manuscript, removing references to Phylotyper identifying novel subtypes and replacing it with the more nuanced description that Phylotyper more frequently properly classifies novel subtypes as a negative result or correctly classifies novel alleles with subtypes in the database. To strengthen this statement, we have as the reviewers suggest, remove entire subtypes from the training set and calculated the false positive rate. Phylotyper was found to have a much lower false positive rate when classifying genes with unknown subtypes.

Phylotyper does not have the ability to create subtype schemes de novo, or to add new subtype classes on the fly. We have not explored this potential application of the Phylotyper method. It would be an interesting avenue to explore as using a phylogenetic model of subtype evolution to define boundaries for new or existing subtypes might make a nice complement for existing de novo stragies like MLST that assign every new allele a separate subtype. However, our goal in this manuscript is to present an alternative genome-based subtype prediction method. De novo subtype creation is outside our intended applications for Phylotyper. Again, we corrected all references to Phylotyper identifing novel subtypes.

6.      “Subtypes are mainly used [..] the main uses of subtype information”. Please revise this paragraph, to allow a better logical flow. Also, Debroy et al have shown the discrepancy between between serotype-based typing and the genetic similarity of O antigen gene clusters, suggesting to replace serotyping by molecular typing. They did not particularly emphasize discrepancies between existing molecular typing scheme and whole genome data as suggested by Whiteside et al.

*Response*:

We have revised this paragraph and clarified our point on the suitability of Phylotyper for predicting subtypes from WGS; namely that Phylotyper helps identify discrepancies between a subtype and genome region that is used for prediction. We cite Debroy et al., as an example of how the associated genetic data and subtypes can have discrepancies.

Minor comments:
1.      The abstract mentions a python package only whereas the text in the implementation mentions python and R. Please revise where necessary.

*Response*:
Fixed.

2.      “The main input into Phylotyper is assembled genome sequences”. Please mention the format type (apparently fasta).

*Response*:
Fixed.

3.      “… in addition to Stx include intimin and serotyping” Please revise the sentence.

*Response*:
Change made.

4.      Replace aslo by also

*Response*:
Fixed.

5.      Revise the sentence “This type of approach, there is…”.

*Response*:
Sentence replaced.

Figure 1.
6.      Please check the label that is cut on the side of the figure. Also, what does the letter at the end of each label correspond to - the expected subtype whereas the predicted subtype is color coded? Please better describe in the legend. Or remove if this is redundant information.

*Response*:
Removed the subtype letters at the end of the label; this was redundant information that was also conveyed by the color. Added information in the figure description on the meaning of the colors. Fixed labels so they fit in the window.

7.      The colors used by Phylotyper could be improved to increase readability of the output for the allele labels. There are several good packages to select colors within R or python.

*Response*:
Switched to discriminatory palette "Set1" in R's RColorBrewer package. Phylotyper uses RColorRamp and this palette to generate sufficient colors for all subtypes.

Sup. Figure 1.
8.      Please plot the legend so that it does not overall the labels

*Response*:
Change made.

9.      Labels are too close and therefore difficult to read if there were several query allele as an input.

*Response*:
Spacing increased.

Supplementary table S2:
1.      Four or five E. coli schemes?

*Response*:
The E. coli serotypes available in Phylotyper comprise two schemes: O- and H-type (here incorrectly counted as one), resulting in five tested schemes. The reference to four has been removed.

Reviewer: 3

Comments to the Author
It's good to see a typing tool that is based on phylogenetic analysis. As public health practitioners still need to compare WGS data to old typing results, this tool will be useful to them.  Moreover, by putting typing into the framework of phylogenetic analysis, this tool can provide an interesting learning/visualization tool for epidemiologists and other public health practitioners to bridge the gap between subtypes (categorical data) and a tree (continuous data). The tool appears to be well implemented.
Some Specific Comments that the authors should address:
Major:
1) Please comment on what happens when an incomplete loci is found in the WGS data. Does the software warn the users of an partial alignment due to incomplete gene?

*Response:
Complete or nearly complete loci sequences are needed for correctly placing the loci in the phylogenetic tree. To prevent incomplete loci being used, we set coverage cutoffs of 90% in the BLAST detection step. No alleles with less than this coverage to one of the reference set sequences will pass the detection step. Users are notified when no loci I have added a description of the requirement for complete loci sequences and the checks we have built into Phylotyper to the Discussion.

2) Please clarify how novel subtypes/alleles are identified and reported.

*Response
Phylotyper is significantly more robust than a direct sequence similarity-based predictor in handling novel subtypes or alleles. We do not mean that Phylotyper performs unsupervised construction of new subtype schemes (although the underlying method could be adapted for this task), rather, that Phylotyper has significantly lower false positive rates when performing prediction on an allele not in the reference set or an allele for which there is no example with the same subtype in the reference set (i.e. novel subtype). In the case of a novel subtype, Phylotyper would not assign the allele any subtype.  Compare this result with a sequence similarity approach, which, in most cases assign the novel subtype the closest subtype based on sequence.

To clarify this distinction, we have added and modified text.  We have also performed and included a subsequent analysis comparing the accuracy of the BLAST-based approach to the Phylotyper approach when the entire subtype group is withheld from the training set to simulate performance when encountering a novel subtype. This analysis shows that Phylotyper is a marked improvement over the sequence-similarity approach, and more frequently returns a negative result.

3) As there is a trend moving to core-genome or whole-genome MLST typing schemes, please comment if Phylotyper could be scaled up to handle that. How would horizontal gene transfer or recombination affect the Phylotyper results?

*Response:
The Phylotyper approach is not likely not applicable to core-genome or whole-genome MLST schemes. The value that Phylotyper provides is that it can take as input gene sequences and return predictions for subtypes based on phenotype or subtypes that cannot be directly inferred from the gene sequences alone. For example, PCR or restriction digestion-based subtyping schemes are an abstraction of the underlying gene sequences and in most cases, cannot be directly inferred from an in silico genome sequence. We have demonstrated the ability of Phylotyper to accurately predict a PCR-based subtype (Stx1 and Stx2) as well as a phenotype-based scheme (O-type:H-type serotype).

Subtypes that are directly identified from the genome sequence do not require Phylotyper's additional prediction step. In MLST, the combination each unique allele sequence defines the type.  Potentially, and this is perhaps what the reviewer is refering to, a specific region of the genome that is smaller than the sequence regions used in an MLST, exists that can be used as a proxy to predict MLST subtype. The benefit in this case, is that the entire genome would not need to be sequenced (or multiple alleles interogated) in order to predict MLST type. There is precedent for this type of approach.  The spa and coa genes in Staphylococcus aureus are used as gene markers to infer phylogenetic and population variation.  The spa and coa types have noted concordance with MLST typing. While this is in interesting potential new application for Phylotyper, it is outside the scope of what we originally designed Phylotyper for.

Horizontal gene transfer or recombination would only impact the result of Phylotyper in situations where the genotype is only correlated with subtype and not causative. If the sequences used to build the subtype scheme produce gene products that determine the subtype, such as the O-antigen processing genes for E. coli being used to predict O-group serotypes, then the prediction would directly be linked to underlying genome. If however unrelated, correlated sequence is used, HGT could disconnect a particular genotype from its subtype. We have added to the manuscript recommendations for users building subtype schemes to focus on genomic regions that have a direct biological role in determining the subtype to reduce possibility of impact by HGT.

Minor:
1) Typo in line 49: also not aslo

*Response*:
Fixed

2) Line 51 mentioned assumptions for building new subtype pipeline.  It'll be good to explain that or if space is an issue, point to the portion of the manual that explains these assumptions

*Response*:
We have added descriptions of these assumputions to the mansucript.

----------------------------------------